---
title: "Human Activity Recognition Modelling"
author: "Kyle Scully"
date: "May 23, 2015"
output: html_document
---
```{r, echo=FALSE}
options(width = 100)
```
# Introduction

# Preparing the Data
### Required Package(s):
```{r warning=FALSE}
library(caret)
library(AppliedPredictiveModeling)
````

## Loading the Data

```{r}
training.raw <- read.csv("./data/pml-training.csv")
testing.raw <- read.csv("./data/pml-testing.csv")
```

```{r}
colnames(training.raw)
```

## Cleaning the Data

First 7 columns do not provide kinametic data so we can remove them as they will not be very good predictors:

```{r}
training.clean <- subset( training.raw, select = -c(1:7) )
testing.clean <- subset ( testing.raw, select = -c(1:7) )
```

## Splitting the Data

We create a validation set by splitting our training set randomly.  Sixty percent (60%) of the original set will be partitioned for training and the remaining forty percent (40%) will be partitioned for our validation set:

```{r}
training.index <- createDataPartition(training.clean$classe, p=0.60, list=F)
training.set <- training.clean[training.index, ]
validation.set <- training.clean[-training.index, ]
```

# Building a Predictive Model

## Machine Learning

## Cross validating the model

# Testing the Final Model

## Appling the model to a test set

### Properly formating the results for submission

### Results

 
 
