---
title: "Human Activity Recognition Modelling"
author: "Kyle Scully"
date: "May 23, 2015"
output: html_document
---
```{r, echo=FALSE}
options(width = 100)
```
# Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

In this project, we used data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants who were asked to perform barbell lifts correctly and incorrectly in 5 different ways.  The data was broken into a training, validation, and testing set.  We then trained a prediction model to classify which of the 5 ways were performed given a random set of kinametic data collected in the same manner.

# Preparing the Data
### Required Package(s):
```{r warning=FALSE}
library(caret)
````
Set the overal seed as many functions will be using random sampling:
```{r}
set.seed(123)
```

## Loading the Data

```{r, cache=TRUE}
training.set <- read.csv("./data/pml-training.csv")
testing.set <- read.csv("./data/pml-testing.csv")
```

```{r, cache=TRUE}
colnames(training.set)
```

## Cleaning the Data

First 7 columns do not provide kinametic data so we can remove them as they will not be very good predictors:

```{r, cache=TRUE}
training.set <- subset( training.set, select = -c(1:7) )
testing.set <- subset ( testing.set, select = -c(1:7) )
```

We remove the columns with NAs as they will likely confound our results:

```{r, cache=TRUE}
training.set <- training.set[, colSums(is.na(training.set)) == 0] 
testing.set <- testing.set[, colSums(is.na(testing.set)) == 0] 
```

We Make sure all values are numeric with the exception of the classe column:
```{r, cache=TRUE}
upper.bound <- which( colnames(training.set)=="classe" ) - 1
training.set <- training.set[, sapply(training.set[1:upper.bound], is.numeric)] 
testing.set <- testing.set[, sapply(testing.set, is.numeric)] 
```

## Splitting the Data

We create a validation set by splitting our training set randomly.  Sixty percent (60%) of the original set will be partitioned for training and the remaining forty percent (40%) will be partitioned for our validation set:

```{r, cache=TRUE}
training.index <- createDataPartition(training.set$classe, p=0.60, list=F)
validation.set <- training.set[-training.index, ]
training.set <- training.set[training.index, ]
```

# Building a Predictive Model

## Machine Learning

We use a random forest algorith with 5 fold cross-validation on our training set:
```{r, cache=TRUE, warning=FALSE}
control <- trainControl(method="cv", 5)
model.randomforest <- train(classe ~ ., data=training.set, method="rf", trControl=control, ntrees=250)
model.randomforest
```

The model seems to have a high accuracy as the estimated out of bag error rate is less than 1%: 

```{r, cache=TRUE}
model.randomforest$finalModel
```

## Applying Model to the validation set
We then apply this model to our validation set to test the accuracy of it on a set seperate from the set it used in training:

```{r cache=TRUE}
predict <- predict(model.randomforest, validation.set)
confusionMatrix <- confusionMatrix(validation.set$classe, predict)
confusionMatrix
```

The accuracy of ~ 99% is plenty sufficent to state that this is a good model for classifying what activity is being performed. Now we can apply the model to our final test set.

# Testing the Final Model

## Appling the model to a test set

We use our model to predict the activities performed in the final test set and then will submitted them online for the assignment to find how accurate the predictions were.

```{r}
testing.prediction <- predict(model.randomforest, testing.set)
testing.prediction
```

### Properly formating the results for submission

```{r}
answers <- testing.prediction
pml_write_files <- function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i], file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```
### Results
All 20 cases in the test set were correctly classified by the prediction model.
 
 
